|

### 2023-06-15 paper/app-eventdata.qmd
finished results part 2

```{r}
#| eval: false
rio::import("../misc/mmp_events_coded_table.csv") %>%
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "",
               caption = "Employed Subset of the MMP Data by Clark and Regan (2019) with Additional and Recoded Variables",
               label = NA) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    latex_options = "repeat_header",
    font_size = 10
  )%>% 
  kableExtra::column_spec(c(1:3, 5), width = "1.8cm") %>%
  kableExtra::column_spec(4, width = "6cm")
```

>>>>

```{r}
rio::import("../misc/mmp_events_coded_table.csv") %>%
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "",
               caption = "Employed Subset of the MMP Data by Clark and Regan (2019) with Additional and Recoded Variables",
               label = NA) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    latex_options = "repeat_header",
    font_size = 10
  )%>% 
  kableExtra::column_spec(c(1:3, 5), width = "1.8cm") %>%
  kableExtra::column_spec(4, width = "6cm")
```

### 2023-06-15 paper/article.qmd
finished results part 2

    [^1]: **Draft version. Please do not circulate or cite without permission.**
subtitle: "Making sense of framing patterns in protest coverage"
format:
  aft-pdf:
    include-in-header:
      - text: |
          \usepackage{siunitx}
  aft-html: default
author:
  - name: Johannes B. Gruber
    affiliations:
      - name: Vrije Universiteit Amsterdam
        department: Department of Communication Science
    orcid: 0000-0001-9177-1772
    email: j.b.gruber@vu.nl
    url: https://www.johannesbgruber.eu/
date: today
filters:
  - abstract-section
keywords: [protest paradigm, media framing, social movements, news reporting, content analysis]
reference-section-title: References
bibliography: references.bib
---

@@

```{r margindframes, fig.cap="Average Marginal Effects for Significant Variables in Individual Frame Models", fig.height=12, fig.width=7}
#| label: fig-margindframes
#| fig-cap: Average Marginal Effects for Significant Variables in Individual Frame Models
#| fig-height: 12
#| fig-width: 7
indfram_marg <- readRDS("../misc/model_margins_ind.Rds")

@@

plot <- indfram_marg %>%
  ggplot(aes(x = AME, y = factor, xmin = lower, xmax = upper)) +
  geom_point(colour = "#939597") +
  geom_pointrange(colour = "#939597") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "#F65058") +
  facet_wrap(~model, ncol = 1, scales = "free_y", strip.position = "top") +
  #facet_grid(rows = vars(model), scales = "free_y", space = "free_y") +
  labs(y = NULL, x = "Average Change in Probability of Frame Usage") +
  theme(strip.background = element_rect(
    fill = "#EEEFF1",
    colour = NA
  ))

@@

Coming soon...

>>>>

    [^1]: **Draft version. Please do not circulate or cite without permission.**
subtitle: "Making sense of framing patterns in protest coverage"
format:
  aft-pdf:
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
  aft-html: default
author:
  - name: Johannes B. Gruber
    affiliations:
      - name: Vrije Universiteit Amsterdam
        department: Department of Communication Science
    orcid: 0000-0001-9177-1772
    email: j.b.gruber@vu.nl
    url: https://www.johannesbgruber.eu/
date: today
filters:
  - abstract-section
keywords: [protest paradigm, media framing, social movements, news reporting, content analysis]
reference-section-title: References
bibliography: references.bib
execute:
  echo: false
  eval: true
---

@@

In the previous section, the theoretical expectations towards what influences the usage of frames were tested.
This was done using grouped dependent variables signifying if one of several delegitimising or legitimising frames were used.
However, the use of this dichotomy has recently been criticised as potentially masking importan mechanisms in protest reporting [@harlowNewParadigm2023].
This section, therefore, discusses regression results for seven different models, each using a different frame as the dependent variable.
The data and model specifications for all of these models were identical to what was discussed above.
Discussion of the results follows the same structure as the previous section using Table 4 and the AME in @fig-margindframes.

```{r}
#| results: asis
indfram_marg <- readRDS("../misc/model_margins_ind.Rds")

@@

dep_ind <- c("Police Violence", "Law & Order", "Cause & Grievances", "Righteous Struggle", 
"Troublemakers", "Decay of Morals", "Nuisance")

models_sel <- models[c("model_logit_1_mlm", "model_logit_2_mlm",
                       "model_logit_3_mlm", "model_logit_4_mlm",
                       "model_logit_5_mlm", "model_logit_6_mlm",
                       "model_logit_12_mlm")]
reorder <- c(2, 5:7, 3:4, 1)

if (is_latex()) {
  texreg(
    models_sel[reorder],
    #custom.model.names = paste0("frame ", c(1:6, 12)),
    custom.model.names = gsub("&", "\\&", dep_ind[reorder], fixed = TRUE),
    custom.header = list("Delegitimising" = 1:4,
                         "Legitimising" = 5:6),
    caption = "Regression Results for Individual Frames",
    custom.coef.names = c(
      "(Intercept)",
      "goal: anti-war ",
      "goal: labour protests",
      "goal: police",
      "goal: social-issue",
      "violent protest",
      "repression of peaceful p.",
      "right-wing",
      "tabloid newspaper",
      "ideological divide: conflict",
      "ideological divide: ambiguous",
      "days from start",
      "year of protest"
    ),
    groups = list("Event-Level Factors" = 2:7, "Outlet-Level Factors" = 8:11, "Time Bound Factors" = 12:13),
    digits = 3,
    stars = c(0.001, 0.01, 0.05, 0.1),
    symbol = "+",
    scalebox = 0.7,
    siunitx = TRUE,
    single.row = FALSE,
    booktabs = TRUE,
    sideways = TRUE,
    label = "tbl-regindframes",
    use.packages = FALSE
  )
} else {
  data.frame(placeholder = "In the PDF, the table is around here") %>%
    custom_df_print(caption = "Regression Results for Individual Frames")
}
```

```{r}
#| label: fig-margindframes
#| fig-cap: Average Marginal Effects for Significant Variables in Individual Frame Models
#| fig-height: 12
#| fig-width: 7
plot <- indfram_marg %>%
  ggplot(aes(x = AME, y = factor, xmin = lower, xmax = upper)) +
  geom_point(colour = "#939597") +
  geom_pointrange(colour = "#939597") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "#F65058") +
  facet_wrap(~model, ncol = 1, scales = "free_y", strip.position = "top") +
  #facet_grid(rows = vars(model), scales = "free_y", space = "free_y") +
  labs(y = NULL, x = "Average Change in Probability of Frame Usage") +
  theme(strip.background = element_rect(
    fill = "#EEEFF1",
    colour = NA
  ))

@@

In the previous section, I concluded that none of the **goals** has a significant effect on the probability that any delegitimising frame is used.
Yet, on an individual frame level, several significant values are observed in Table 4.
The Nuisance frame is significantly more likely to be utilized in the case of labour protests than for the control category (AME: `r marg_ind("goal: labour protests", model = "Nuisance")`).

For protests against police brutality, the effects on various frames counteract each other.
It's considerably more likely that Law & Order (AME: `r marg_ind("goal: police", model = "Law & Order")`) and the Troublemakers (AME: `r marg_ind("goal: police", model = "Troublemakers")`) frames are used in the coverage of police protests.
Conversely, it is less likely that the Decay of Morals (AME: `r marg_ind("goal: police", model = "Decay of Morals", abs = FALSE)`) and/or the Nuisance (AME: `r marg_ind("goal: police", model = "Nuisance", abs = FALSE)`) frame are utilized.
This means some delegitimising frames are more probable while others are less so, causing the likelihood of any delegitimizing frame being used to remain roughly the same.

The chances of the Law & Order frame being used, revealing the positive influence of police presence, increases for protests against police brutality.
However, a higher AME is observed in the model for the antithetical frame, Police Violence (AME: `r marg_ind("goal: police", unit = "percent", model = "Police Violence")`).
That means that protests against police brutality are divisive:
they lead to more coverage highlighting claims of police brutality and providing background on them;
but coverage will also defend police actions.
They lead to more coverage emphasizing claims of police brutality while also defending police actions.

The final goal, social issues, only significantly alters the result in the model for the Law & Order frame.
Specifically, it's `r marg_ind("goal: social-issue", unit = "percent", model = "Law & Order")` less probable that the Law & Order frame is utilized for social-issue protests in comparison to political protests.

In the model for aggregated legitimizing frames, social issue and labour protests were given preference.
Here we see that this effect is solely due to the increased likelihood of *Cause & Grievances* being used (AME: `r marg_ind("goal: labour protests", model = "Cause & Grievances", abs = FALSE)`).
While the cause & grievances frame is significantly more probable (AME: r marg("goal: social-issue", model = "Cause & Grievances", abs = FALSE)), the righteous struggle model doesn't show any significant effect.

The remaining frame, police violence, neither legitimising nor delegitimising, displays the most obvious evidence that different goals influence the journalists' choice to employ a frame.
Anti-war and political protests use this frame in the same manner.
All other demands lead to significantly different uses of this frame.
Labour and social-issue protests are less likely to resort to police violence (AME: `r marg_ind("goal: labour protests", model = "Police Violence", abs = FALSE)`, `r marg_ind("goal: social-issue", model = "Police Violence", abs = FALSE)`), whereas anti police brutality protests are significantly more likely to be covered with this frame (AME: `r marg_ind("goal: police", model = "Police Violence", abs = FALSE)`).

The outcomes for individual frames corroborate the theoretical idea behind H1a-b:
the goal of a protest matters for how it's covered.
However, results are more complex than anticipated.
Not all protest events with certain types of goals are covered with frames that are more or less favourable for protesters.
Instead, different protest demands seem to steer the focus of reporting.

When it comes to protest **tactics**, the models for individual frames suggest that the positive impact on delegitimising framing was propelled by the use of the *Law & Order* (AME: `r marg_ind("violent protest", model = "Law & Order", abs = FALSE)`) and *troublemakers* (AME: `r marg_ind("violent protest", model = "Troublemakers", abs = FALSE)`) frames.
The other two delegitimising frames actually have significant negative effects, although of a lesser magnitude.
This is logical as both *law & order* and *troublemakers* are primarily centered around violence.

*Decay of morals* and *nuisance*, on the other hand, are less likely used respectively by `r marg_ind("violent protest", model = "Decay of Morals", unit = "percent")` and `r marg_ind("violent protest", model = "Nuisance", unit = "percent")` when violence is present.
In such scenarios, journalists who wish to highlight the negative aspects of an event likely opt for a frame that directly focuses on violence.
This notion is supported by the *police violence* model, which indicates that journalists tend to exercise a healthy amount of scepticism when violence unfolds during a protest.
Although AMEs are not directly comparable across models, it's evident that the average change of probability, `r marg_ind("violent protest", model = "Police Violence", unit = "points")` percentage points, is greater than in the models for both *Law & Order* and *Troublemakers*. 
Therefore, in this aspect, the model for the use of *Police Violence* is more akin to those for delegitimising frames.

Table 4 further reveals that the negative effect for violence seen in the legitimising model was entirely driven by *cause & grievances*.
As violence occurs, journalists are less likely to follow the reasoning of protesters, a finding which basically supports what is already common knowledge.
Interestingly, the model for *righteous struggle* is the only one that shows no significant variance when violence is involved in an event.
This implies that journalists depict a cause as just regardless of violence occurrence, which seems somewhat hard to believe.
However, remember that the use of this frame is also independent of other frames' use.
In other words, even when journalists acknowledge a cause as just, the occurrence of violence still prompts them to concentrate on tactics over the event's message.

The models for individual frames present a somewhat perplexing scenario when it comes to the variable *state repression of peaceful protest*.
Despite the delegitimising frames model indicating no effect, three out of the four frames in the category do exhibit significant values.
However, the positive probabilities for *law & order* and *nuisance* seem to be offset by the negative effect for *decay of morals*.

This suggests that while it is `r marg_ind("repression of peaceful p.", model = "Law & Order", unit = "percent")` more likely that *law & order* is used, and `r marg_ind("repression of peaceful p.", model = "Nuisance", unit = "percent")` more likely that *nuisance* is used, the overall delegitimising framing remains stable when protestors are *arrested*, as *decay of morals* is used significantly less frequently (AME: `r marg_ind("violent protest", model = "Decay of Morals")`).
It appears that when arrests occur, trouble is emphasized, but as the *troublemakers* model shows no significant impact, protestors are not directly held accountable for it.

It's noteworthy that a negative effect exists for *state repression of peaceful protest* in the *decay of morals* model.
This frame underscores the risk of protestors setting a bad example for others and disrupting the general political consensus.
One potential explanation could be that journalists favor using a frame that accentuates trouble when protestors are arrested, believing that it adequately covers what they would otherwise mention about morals.
However, the presence of a significant positive effect on the usage of the *nuisance* frame contradicts this hypothesis.

In the two models for legitimising frames, the narrative from the aggregated model is echoed:
there's a significant negative effect on the likelihood of coverage using either *cause & grievances* or *righteous struggle*.
Theoretical expectations of what happens when there's *state repression of peaceful protest* fit even less well for the models of individual frames. 
The rule seems to be that getting arrested, even during peaceful demonstrations, leads to reduced probabilities of coverage that would benefit a protest.
The sole exception to this rule is the *decay of morals* frame, which is used less frequently in these instances.

On the outlet level, one of the main questions was if a *newspaper's ideology* or the distance between the newspaper's and protesters' ideology, i.e., the *ideological divide*, matters for reporting --- with the surprising result that neither of them does.
This is only slightly different for individual frames.
There is a significant negative effect in the models for *righteous struggle*:
for articles published in right-leaning outlets, the probability that this frame is used is `r marg_ind("right-wing", model = "Righteous Struggle", unit = "points")` percentage points lower compared to left-leaning outlets.
This result is little surprising as *righteous struggle* basically highlights the advantages protest has for an issue.
More surprising is the complete lack of other significant effects of this variable in the other models.
There is an effect for *law & order*, suggesting that the frame is *less* common in right-leaning outlets.
Alas, the p-value is 0.09, which is usually not considered significant.
Overall then, it seems that *newspapers' ideologies* play a small role when it comes to the framing of protest events.

As in the aggregated models, there is no support for the two hypotheses regarding the impact of an *ideological divide* between protest and reporting newspaper.
Although all of the models for delegitimising frames show p-values just over the significance threshold of $p < 0.05$.
And defying theoretical expectations again, the *police violence* frame, which highlights the responsibility of the police for clashes, is `r marg_ind("ideological divide: conflict", model = "Police Violence", unit = "percent")` more likely when there is an ideological divide.
The impression from the last section that the impact of newspaper ideology on protest coverage is incidental instead of systematic is further reinforced by these findings.

Only two of the seven models for the usage of individual frames show any significant effect for **type of newspaper**:
*righteous struggle* is used `r marg_ind("tabloid newspaper", model = "Righteous Struggle", unit = "percent")` less likely by tabloid newspaper, *police violence* is used `r marg_ind("tabloid newspaper", model = "Police Violence", unit = "percent")` more likely.
This means that **H6b** was accepted because *righteous struggle* is used more likely in quality media, while the newspaper type has no significant effect on the other legitimising frame.
All delegitimising frames are used regardless of the type of newspaper, which confirms the finding from the last section.

In terms of **temporal factors**, my findings for the aggregated frames suggest that the *days since the start* of a protest only significantly influence delegitimising frames. 
This story is repeated in the models for individual frames, providing more evidence that the theory of a *news framing cycle* [@gottliebProtestNewsFraming2015], does not hold up for legitimising framing in the UK:
instead of becoming more common after the first phase of a protest, *righteous struggle* is less likely to be used over time, while there is no effect for *cause & grievances*.

For two of the four delegitimising frames, however, the theoretical expectation is further supported:
they are less likely to be used the longer a protest is in the past.
However, there is no effect for *nuisance* and a positive effect for *decay of morals*.
This fits @gottliebProtestNewsFraming2015’s argument who suggests not that specifically delegitimising frames are used in the early stages of protest coverage, but that "conflict frames" (p. 7) are employed most often.
When conflict itself becomes less interesting, it appears that *decay of morals*, which highlights that protesters are a fringe group and set a bad example, gains in importance.

The models for individual frames also provide further evidence that legitimising frames gain in importance over the years.
Compared to 1992, the use of *cause & grievances* is `r marg_ind("Year of protest (since 1992)", model = "Cause & Grievances", unit = "percent")` more likely in 2017.
The only delegitimising frame that is influenced by a temporal shift is *law & order*, which becomes less likely to be used as time goes on (AME: `r marg_ind("Year of protest (since 1992)", model = "Law & Order", abs = FALSE)`).
Interestingly, *Police Violence*, again, aligns with the trend of the legitimising frames, as Table 4 reveals a significant positive effect for the *year* variable (AME: `r marg_ind("Year of protest (since 1992)", model = "Police Violence", abs = FALSE)`).
